\documentclass[11pt]{article}
\input{../Bibliography/commoncommands}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.0in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{18pt}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.25in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parindent}{0.0in}
\setlength{\paperheight}{11.0in}
\setlength{\paperwidth}{8.5in}
\setlength{\parskip}{.5\baselineskip}

\newcommand{\paper}{note}
\newcommand{\citesmv}{\cite{Smokeview_Users_Guide}}

\begin{document}

\bibliographystyle{unsrt}


\pagestyle{plain}

%\doublespace

\pagestyle{empty}
%\tableofcontents
%\listoffigures
%
% ----------------------  first cover/title page --------------------------
%

\pagenumbering{arabic}
\newpage

PROCEEDINGS, Fire and Evacuation Modeling Technical Conference 2018 \\
Gaithersburg, Maryland, October 1-3, 2018

\begin{center}
{\bf VISUALIZING SMOKE AND FIRE} \\
\hspace{1in} \\
Glenn Forney \\
National Institute of Standards and Technology \\
Gaithersburg, Maryland, USA \\
e-mail: glenn.forney@nist.gov \\
\hspace{1in} \\
\end{center}
\begin{center}
{\bf ABSTRACT}\\
\end{center}

This note will discuss some of the physics and associated numerical algorithms used by Smokeview to visualize smoke and fire. Recent progress accounting for emission from CO2 will also be discussed.  Realistic visualization methods are important for applications where one wishes to observe qualitative effects of fire and smoke rather than quantitative characteristics such as temperature or velocity. This would be the case for a fire fighter using a computer based fire fighting simulator.  Smoke visualization methods considering interactions between light and smoke, require the solution of the radiation transport equation (RTE) also called the volume rendering equation. This equation models how light is affected by its interaction with smoke, a participating medium. Smokeview uses the RTE to account for extinction (absorption plus out-scattering) by the smoke and emission from the fire. The form of the RTE used by Smokeview to model smoke and fire appearance is identical to that used by FDS to model radiative heat transfer.  The implementation details are different however.  Smokeview uses an extinction coefficient appropriate for visible light while FDS use one appropriate for infrared. Smokeview requires a solution at only one position, the observer's viewpoint, while FDS requires a solution at many points within the solution domain. Even so, the solution of these equations for visualization requires significant computational and memory resources.  Approximations are then required to display smoke and fire at interactive frame rates. The primary approximation is to take advantage of the low albedo character of smoke allowing one to either simplify or eliminate scattering terms in the RTE.

%
% -------------------  Introduction ------------------------
%

\input{smoke_notes2}

%
% -------------------  References ------------------------
%
\newpage
\bibliography{../../../fds/Manuals/Bibliography/FDS_general,../../../fds/Manuals/Bibliography/FDS_refs,../../../fds/Manuals/Bibliography/FDS_mathcomp,../Bibliography/sv_fire,../Bibliography/sv_graphics}

\addcontentsline{toc}{chapter}{Bibliography}

\end{document}
